---
title: "HW 4"
author: "General Linear Models 1"
date: "Due Date: 10/02/2020"
output: pdf_document
---

# Data for this Assignment
Data for assignment come from the following manuscript: 
Shaffer, V. A., Scherer, L. D., Focella, E. S., Hinnant, A., Len-Rios, M. E., Zikmund-Fisher, B. J. (2017). What is the story with narratives? How using personal narratives in journalism changes health behavior. Health Communication, doi: http://dx.doi.org/10.1080/10410236.2017.1333562 


This study was designed to examine the effect of a health news story about a person who developed a life-threatening illness (Stevens-Johnson Syndrome, SJS) after using Advil on subsequent Advil use (or other medication with ibuprofen as its active ingredient). All participants completed measures of numeracy (Subjective Numeracy—Fagerlin et al., 2008; Berlin Numeracy Test—Cokely, Galesic, Schulz et al., 2012), health literacy (Chew, Bradley, & Boyko, 2004), and thinking style (Rational-Experiential Inventory, 24-item short form). 

The data file can be found in a .csv file on Canvas. The file contains data on 7 variables from 443 participants. Below are the definitions of these measures. Note the data file contains missing data labeled ‘NA’.

# Items in the data file
* Likelihood_TIME1—‘How likely are you to use Advil (or other medication with ibuprofen as the active ingredient) in the future?’ 1-5 point Likert scale
*	Likelihood_TIME2—‘How likely are you to use Advil (or other medication with ibuprofen as the active ingredient) in the future?’ Assessed at two week follow up; same scale as above  
* Rational_subcale—Rational subscale of the REI (Rational Experiential Inventory)
* Experiential_subscale—Experiential subscale of the REI
* HealthLit_SUM—Health literacy screening scale
* SNS—Subjective Numeracy Scale
* Berlin_cat—Berlin Numeracy Scale

# Instructions for this assignment
Use R Markdown to create a pdf answer sheet for this assignment, showing your code for each item as appropriate. Do all items below using R commands (unless you choose to use GPower for 7 and 8, in which case simply type your answers from GPower into your R Markdown document). 

In this assignment, you will be testing the hypothesis that the Rational subscale of the REI is a significant predictor of intentions to use Advil in the future measured at Time 2 (two-weeks after the intervention)—**centering the predictor at the mean**

# Questions

1. First, write out the equations for Model A and Model C that test this hypothesis (2 points)

**Model Comparison**
\begin{align*}
\text{Model A (Augmented Model): } Y_i &= \beta_0 + \beta_1(X_i-\overline{X}) + \epsilon_i \\
\text{Model C (Compact Model): } Y_i &= \beta_0 + \epsilon_i \\
\end{align*}

```{r}
HW4 <- read.csv("~/Dropbox/PSYCH 8710 GLM I/Datasets/Psych Faculty datasets/Shaffer Lab/Shaffer Lab Data.csv")
```

2. Create descriptive summaries of the predictor and outcome variables. (1 point)
```{r}
summary(HW4$Rational_subscale)
summary(HW4$Likelihood_TIME2)
```

3. Construct a scatterplot showing the bivariate relationship between the predictor and the outcome. Use the "jitter" command to better see all of the observations on the Likert-scale item.(1 points)

```{r}
plot(HW4$Rational_subscale,jitter(HW4$Likelihood_TIME2,2), 
     xlab = "Rational Subscale", ylab = "Likelihood of Ibuprofen Use", main= "Shaffer Lab Data")
```

4. Using the lm command, test the hypothesis that the Rational subscale of the REI is a significant predictor of intentions to use Advil in the future measured at Time 2 (two-weeks after the intervention)—**centering the predictor at the mean**. (2 points)

```{r}
Rat_mean <- mean(HW4$Rational_subscale, na.rm = TRUE)
HW4$Rational_meancent <- HW4$Rational_subscale-Rat_mean 
ModelA <- lm(Likelihood_TIME2 ~ Rational_meancent, data = HW4)
summary(ModelA)
```

5. Write an interpretation for the intercept and slope of Model A (2 points)
* Intercept: the predicted value of likeihood of using ibuprofen (3.53) when the Rational subscale score is zero. Because the predictor is mean-centered, 3.53 is also the mean of Y.
* Slope: the predicted increase (0.03) in likelihood of using ibuprofen for every 1-unit increase in the Rational subscale.

6. Calculate a measure of effect size for this relationship between X (Rational subscale of REI) and Y (Likelihood of ibuprofen use at Time 2). Interpret this value (2 points)
```{r}
eta_squared <- 1-((1-.022)*((261-1)/(261-2)))
eta_squared
```

* PRE is given in output (.02). The inclusion of the rational subscale as a conditional predictor of likelihood of ibuprofen use reduce model error by 2%. 
* Could alternatively calculate the unbiased estimate of eta squred, which would tell us similarly that this is a small effect size.

7. Calculate the power associated with the test of the Rational subscale as a predictor of behavioral intentions (1 point)
*This can be done using G-power. Doing this in R would also be acceptable, we just haven't formally taught this yet in class.
*In G-Power, you would get an observed power of 0.68 (need to calculate the standard deviation of X and Y, which is why I have this computed below)
```{r}
sdRat <- sd(HW4$Rational_subscale, na.rm=TRUE)
sdLik <- sd(HW4$Likelihood_TIME2, na.rm = TRUE)
sdRat
sdLik
```

8. Imagine we wanted to replicate this test with 80% power. What sample size is needed for this replication? (1 point).

* Using G-Power, this sample size needed to replicate is 348.

9. Examine the data for outliers in the predictor variable. Which (if any) observations would you consider as potential outliers? (2 points)

```{r}
#Calculating leverage
library(olsrr)
lev = hat(model.matrix(ModelA))
plot(lev)
HW4[lev>0.03,]
```

* I do not see any values of leverage that are so extreme that I would consider calling them outliers. 

10. Examine the data for outliers in the outcome variable.Which (if any) observations would you consider as potential outliers? (2 points)

```{r}
#Calculating Studentized Deleted Residuals
library(olsrr)
ols_plot_resid_stud(ModelA)

#HW4$stud.del.resid <- rstudent(ModelA, na.rm = TRUE)
#ols_plot_resid_stud(ModelA)
#HW4$r<-rstudent(ModelA)
```

* I do not see any values of studentized deleted residuals greater than 3. Therefore, I would not conisder any of the observations outliers based on this plot.

11. Examine the data for outliers in the joint effect of the predictor and the outcome variable. Which (if any) observations would you consider as potential outliers? (2 points)

```{r}
#HW4$CooksD <- cooks.distance(ModelA)
#plot(HW4$CooksD, ylab = "Cooks Distances")
ols_plot_cooksd_bar(ModelA)
ols_plot_cooksd_chart(ModelA)
```

* I do not see any values of Cooks D that are so extreme that I would consider calling them outliers.

12. Examine the residuals for systematic violations of the assumption that the errors are normally distributed. How would you interpret this? (2 points)
```{r}
qqnorm(ModelA$residuals)
qqline(ModelA$residuals)
hist(ModelA$residuals)
```

* Although the data appear to have thinner tails than the normal distribution, I would not be concerned about violating the assumption of normality given this assumption is very robust to this type of violation.

13. Examine the data for systematic violations of the assumption that the data are identically distributed. How would you interpret this? (2 points)
```{r}
library(car)
plot(ModelA)
```

* Looking at the scale location plot, I do not see any evidence of heteroscedasticity.







